You are the way you (structurally) talk: Structural-temporal neighbourhoods of posts to characterize users in online forums

### 2016-05-05 doc/2016-05 Luchon /presentation_SIG_Toulouse_Avril_2016.tex

\hypersetup{colorlinks,urlcolor=blue}
\AtBeginSection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection]
		\end{NoHyper}
	\end{frame}
}
\begin{document}
\addtobeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\title{You are the way you (structurally) talk:  Structural-temporal neighbourhoods of posts to characterize users in online forums}
\author{Alberto Lumbreras \\Jouve B., Velcin J., Guégan, M.}
\date{April 8, 2016}
\maketitle

@@

	\vfill
	Extract forum of interest:\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}science}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}france}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}sociology}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}complexsystems}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}podemos}} $\leftarrow$ in this presentation\\
	...
\end{frame}

@@

\begin{frame}{Graph representations}{Which one is better?}

\begin{itemize}
\item Depends on the task!
\item One might choose multiple representations (multi-level analysis)
\end{itemize}
\vfill
My choice: 
\begin{itemize}
\item Mostly tree representation
\item Because it explicitly represents discussions (and their evolution).
\end{itemize}
\begin{figure}
	\centering
	\includegraphics[width=0.3\textwidth]{tree1}	
	\includegraphics[width=0.3\textwidth]{tree2}	
	\includegraphics[width=0.3\textwidth]{tree4}	
\end{figure}
And sometimes:
\begin{itemize}
\item SNA representation of single conversations.
\end{itemize}
\end{frame}

\section{Structures of conversations}
\subsection{Basic idea}
\begin{frame}{Intuition}
\textit{Hypothesis}: different individuals have tendency towards different types of conversations and these types are reflected in the structure of their interactions.
\vfill
These conversational structures might be observed at two levels (at least):
\begin{itemize}
\item Social graph.
\item Posts graph (tree)
\end{itemize}
\end{frame}

@@

\begin{frame}{Order-based neighbourhoods}{Census}
We found 129 different motifs:
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{neighbourhoods_2_4_1}
	\end{figure}
\end{frame}

\begin{frame}{Time-based neighbourhoods}{Definition}
\begin{itemize}
	\item 1. Extract neighbourhood of post $i$ with radius $r$.
	\item 2. Detect changes of speed (vertical/horizontal changepoints)
	\item 3. From $i$, get the posts around until a changepoint is found.
\end{itemize}
%\vspace{cm}  
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{breakpoints2}
	\end{figure}
\end{frame}

\begin{frame}{Time-based neighbourhoods}{Census}
We found 165 different motifs:
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{neighbourhoods_time_1}
	\end{figure}
\end{frame}

@@

\begin{frame}{Conversation-based clustering}{Order-based}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{PCA_cluster_order_based}
	\end{figure}	
\end{frame}

\section{Conclusions}
\begin{frame}{Conclusions}
	\begin{itemize}
	\item \textbf{Q: Can we use graph structure to characterise users?}
	\item A: Yes!
	\end{itemize}	
	\begin{itemize}
	\item \textbf{Q: By using triads}?
	\item A: No. They are not useful in trees.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: So, what kind of structure?}
	\item A: Posts neighbourhoods that are time/order sensitive.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: What about language?}
	\item A: It's ok, but structure is more directly linked to thread dynamics (future work)
	\end{itemize}
	\textbf{Future work:}
	\begin{itemize}
		\item Prune time-based neighbourhoods to reduce dimensionality.
		\item Do users jump from cluster to cluster (paths of roles)
	\end{itemize}	
\end{frame}

>>>>

\hypersetup{colorlinks,urlcolor=blue}
\AtBeginSection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection]
		\end{NoHyper}
	\end{frame}
}
\AtBeginSubsection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection, currentsubsection]
		\end{NoHyper}
	\end{frame}
}
\AtBeginSubsubsection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection, currentsusubbsection]
		\end{NoHyper}
	\end{frame}
}
\begin{document}

\addtobeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\title{You are the way you (structurally) talk:  Structural-temporal neighbourhoods of posts to characterize users in online forums}
\author{Alberto Lumbreras \\Jouve B., Velcin J., Guégan, M.}
\date{May 3, 2016}
\maketitle

@@

	\vfill
	Extract forum of interest:\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}science}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}france}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}sociology}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}complexsystems}}\\
	\href{www.reddit.com/r/gameofthrones}{www.reddit.com/r/{\color{red}gameofthrones}} $\leftarrow$ in this presentation\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}podemos}} $\leftarrow$ in this presentation\\
	...
\end{frame}

@@

\begin{frame}{Graph representations}{Conversations are trees}

\begin{itemize}
\item Explicit structure.
\item Dynamic (order, time)
\end{itemize}
\begin{figure}
	\centering
	\includegraphics[width=0.3\textwidth]{tree1}	
	\includegraphics[width=0.3\textwidth]{tree2}	
	\includegraphics[width=0.3\textwidth]{tree4}	
\end{figure}
\end{frame}

\section{Structures of conversations}
\subsection{Basic idea}
\begin{frame}{Intuition}
\textit{Hypothesis}: different individuals have tendency towards different types of conversations and these types are reflected in the structure of their interactions.

\end{frame}

@@

\begin{frame}{Time-based neighbourhoods}{Definition}
\begin{itemize}
	\item 1. Extract neighbourhood of post $i$ with radius $r$.
	\item 2. Detect changes of speed (vertical/horizontal \textit{changepoints}) (PELT algo)
	\item 3. From $i$, get the posts around until a changepoint is found.
\end{itemize}
%\vspace{cm}  
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{breakpoints2}
	\end{figure}
\end{frame}

\begin{frame}{Colouring}
Label special nodes:
\begin{itemize}
	\item Red: ego.
	\item Yellow: parent of ego (and posts of same author)
	\item Orange: other posts by ego author
	\item White: root
\end{itemize}
	\begin{figure}
		\centering
		\includegraphics[width=0.3\textwidth]{colouring}
	\end{figure}

\end{frame}
\begin{frame}{Some frequent neighbourhoods}{}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{dictionary1}
	\end{figure}
\end{frame}
\begin{frame}{Some frequent neighbourhoods}{}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{dictionary2}
	\end{figure}
\end{frame}

\subsection{Comparing neighbourhoods}
\begin{frame}{Frequency distribution}{}
	\begin{figure}
		\centering
		\includegraphics[width=0.8\textwidth]{compare_census_distribution}
	\end{figure}
Time-based (black) reduces the space w.r.t structure-based (blue)
\end{frame}

\begin{frame}{Frequency distribution}{}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{compare_census_distribution_cum}
	\end{figure}
\end{frame}

\begin{frame}{Discrepancies}{Time-based vs Order-based neighbourhood}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{confusion_time_order}
	\end{figure}
\end{frame}

\begin{frame}{Discrepancies}{Basic neighbourhood vs Time-based neighbourhood}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{confusion_struct_time}
	\end{figure}
\end{frame}

\begin{frame}{Structure-based vs Order-based vs Time-based}
Structure-based:
\begin{itemize}
\item  too big (and too many) neighbourhoods.
\end{itemize}
Order-based:
\begin{itemize}
\item  Dominance of monoid hides richer conversational structures.
\end{itemize}
Time-based:
\begin{itemize}
\item Space more reduced than simple structure-based.
\item Criteria to choose radius dynamically (r = until conversation slows down)
\end{itemize}
\end{frame}

@@

\begin{frame}{Conversation-based clustering}{Time-based}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{cluster_whiskers}
	\end{figure}	
\end{frame}

\begin{frame}{Conversation-based clustering}{Interpretation}
	Greens:
	\begin{figure}
		\centering
		\includegraphics[height=0.17\textwidth]{motif8}
		\includegraphics[height=0.17\textwidth]{motif4}
		\includegraphics[height=0.17\textwidth]{motif15}
	\end{figure}	
	Reds:
	\begin{figure}
		\centering
		\includegraphics[height=0.17\textwidth]{motif20}
		\includegraphics[height=0.17\textwidth]{motif3}
		\includegraphics[height=0.17\textwidth]{motif19}	
		\includegraphics[height=0.17\textwidth]{motif14}	
		\includegraphics[height=0.17\textwidth]{motif13}	
		\includegraphics[height=0.17\textwidth]{motif18}	
	\end{figure}	
	Blues (avoid these motifs):
	\begin{figure}
		\centering
		\includegraphics[height=0.17\textwidth]{motif3}
		\includegraphics[height=0.17\textwidth]{motif19}
	\end{figure}	
	\end{frame}

\section{Conclusions}
\begin{frame}{Conclusions}
	\begin{itemize}
	\item \textbf{Q: Can we use graph structure to characterise users?}
	\item A: Yes!
	\end{itemize}	
	\begin{itemize}
	\item \textbf{Q: By using triads}?
	\item A: No. They are not useful in trees.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: So, what kind of structure?}
	\item A: Posts neighbourhoods that are time/order sensitive.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: What about language?}
	\item A: It's ok, but structure is more directly linked to thread dynamics (future work)
	\end{itemize}
\end{frame}
